---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Load in the FAS Litters Data

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## 'select'

choose some columns and not others

What we wanna keep:

```{r}
# we select 2 columns 'group' and 'litter_number'.
select(litters_df, group, litter_number) 
```

```{r}
# select a range of column 'from gd0_weight to gd_of_birth' and the column 'group'
select(litters_df, group, gd0_weight:gd_of_birth)
```

What we wnna lose:

```{r}
# this only get rid of the column 'litter_number'
select(litters_df, -litter_number)
```

Renaming columns while selecting...

```{r}
select(litters_df, GROUP = group, LITTer_NUmBER = litter_number)
```

Rename only, keep all the var. columns

```{r}
rename(litters_df, GROUP = group, LITTer_NUmBER = litter_number)
```

Select helper:
?select_helpers (type in the console)

```{r}
#select everything that starts with gd
select(litters_df, starts_with("gd"))
```

```{r}
# move the litter_number to the first column, and then the rest of columns
select(litters_df, litter_number, everything())
```

```{r}
# reloate() does the same thing above, move litter_number to the beginning
relocate(litters_df, litter_number)
```

## filter
It's intended to remove rows or keep rows that u are interested in retainning for some subsequent analysis.So select every single row every time we did, it just reduced or changed or reorganized the columns. Select operates on columns filter now is gonna operate on rows.

```{r}
filter(litters_df, gd0_weight < 22)
# we get only 8 obs out of the original 48 rows for which gestational day zer weight is less than 22. It's kept all my other columns.

```

```{r}
filter(litters_df, gd0_weight >= 22) #26 rows

```

```{r}
filter(litters_df, gd_of_birth == 20)
```

```{r}
filter(litters_df, !(gd_of_birth == 20))
filter(litters_df, gd_of_birth != 20) #these 2 are the same
```

```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20) # select obs based on 2 conditions 
```

```{r}
filter(litters_df, group == "Mod8") #for character var
```

```{r}
filter(litters_df, group %in% c("Mod8", "Con7")) 
# use %in%, select obs based on variable 'group' where group is equal either to Mod8 or Con7
```


## mutate
It allows us either to modify an exising variable or create some new variable

```{r}
mutate(
  litters_df, 
    wt_gain = gd18_weight - gd0_weight, 
    group = str_to_lower(group))
  
  
```




