---
title: "Data Import"
output: github_document
---

# we focus on readr package.

```{r setup}
library(tidyverse)
library(readxl) # for other file format: excel
library(haven) # for other file format: SAS

```

## Read in some data

Read in the litters dataset.

```{r}
litters_df = read_csv("./data/FAS_litters.csv") 
# read_csv is a fucntion in reader package
# give a relative path "./data/FAS_litters.csv"
# starting here "data_wrangling_i" go into the data folder and go into fas litters.csv
# absolute path would be starting from the really begin of your computer, says the desktop

litters_df # take a look

# overwrite the original dataset: clean_names() with a transformed version
# janitor::clean_names() means in the janitor package, use function clean_names(); or we can load the package by library(janitor), but we're not, b/c clean_names() is the only function we use.
litters_df = janitor::clean_names(litters_df)

```

## Take a look at the data

Printing in the console

```{r}
litters_df  # variables names are overwritten nicely by snake case with no special characters

```

```{r}
head(litters_df)
```

```{r}
tail(litters_df) # check the tail of the dataset
```

```{r}
skimr::skim(litters_df) # we wanna use the skim() function inside of the package "skimr"
# this is a little bit gimmicky, b/c it's sth cool when u check it out and then u almost end up never use it in practice
```

in the console, we type View(litters_df) to open a new window to look at the full dataset.

dbl means double, it's a particular kind of number in cs and programming

## Options to read_csv
Additional options to the read_csv() function.
?read_csv in the console to look at the help file on the right.

```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE) 
# skip = 10, means skip the first 10 rows
# col_names = FALSE, means make up column names as X1 to X8 
# na = c("", "NA", ".", 999) means anytime if u see these, treat as a missing value
# col_type
```

## Other file formats

##### Read in an excel file.

```{r}
#read_excel() in the package "readxl"
mlb_df = read_excel("./data/mlb11.xlsx", range  = "A1:F7") 
mlb_df
```

In the console, type:
View(mlb_df) # to look the full data set
?read_excel # the help file

##### Read in a SAS file

```{r}
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```

## Comarison with Base R

what about 'read.csv' ....?

```{r}
litters_base = read.csv("data/FAS_litters.csv")
litters_readr = read_csv("data/FAS_litters.csv")

litters_base # by default print the entire data, but it'smore difficult to look at
litters_readr # this is better

```

## Exporting data

Export the mlb sub-table.

```{r}
write_csv(mlb_df,"./data/mlb_subtable.csv") 
# we export the tables from mlb_df to the folder "data".
```


